RCloud: DevOps for Visual Analysis and Data Science

### 2014-03-18 paper.tex
yes

\title{RCloud: Integrating Exploratory Visualization, Analysis and Deployment}

% DevOps for Visual Analysis and Data Science

@@

%% Abstract section.
\abstract{ Consider the emerging role of a data science team within an
  organization. Individual data scientists and statisticians usually
  work on loosely related problems, and must devise effective ways to 
  share their findings and to move results from exploratory data analysis
  to automated diagnostics and reports deployed for wider consumption.
  There are two problems with the current
  practice. First, there are gaps in this workflow: EDA is performed
  with one set of tools, and automated reports and deployments with another.
  Second, EDA environments often assume a single-developer perspective,
  while data scientist teams could greatly benefit from easier sharing of
  scripts and data feeds, experiments, annotations, and automated
  recommendations, which are well beyond what traditional version control
  systems provide.  \todo{sell idea, not system} In this paper, we propose RCloud, a system that supports
  collaborative data analysis, visualization and web deployment. We will
  discuss the design decisions, tradeoffs and limitations, and compare RCloud
  to other current proposals.
} % end of abstract

@@

% What's the area
Consider the emerging role of a data science team within an
organization \cite{Keim:2008:VAS}.

% How do people do it today?
Individual data scientists and statisticians usually
work on loosely related problems, and must devise effective ways to 
share their findings and to move results from exploratory data analysis
to automated diagnostics and reports deployed for wider consumption.
They generally rely on a patchwork of resources that are shared informally,
or at best, through systems that attack some of the tasks or functions in
the data analysis workcycle.

% Why does this not work?
There are two problems with the current
practice. First, there are gaps in this workflow:
Exploratory Data Analysis (EDA) is performed with one set of tools,
and automated reports and deployments with another.
Second, EDA environments often assume a single-developer perspective,
while data scientist teams could greatly benefit from easier sharing of
scripts, data sets, data feeds, experiments, annotations, and automated
recommendations, which are well beyond what traditional version control systems provide.

% What are we going to do about this?
\todo{Sell idea, not system} In this paper, we investigate the hypothesis
that the practice of visual analytics can advance by adopting techniques
from information retrieval and collaborative computer, formally supported
within the computing environment. To investigate this, we propose RCloud,
a system that supports collaborative data analysis, visualization and
web deployment. We will discuss the design decisions, tradeoffs and
limitations, and compare RCloud to other current proposals.

\todo{Define the requirements for the system}
\begin{itemize}
\item Technology transfer. Things developed by analysts need to become accessible in the production environment and should be supported in the environment, making it easy to push new features, experiments, etc.
\item Dasboarding. There should be a mode in which people can operate the experiments without coding. At times code has to be written because some things cannot be done any other way but other people must not ever see code because their eyes will fall out.
\item Version control as a primary concept in the repository. Need to control the visibility of 
\item Modern interactive technologies. People want to interact with data to explore it. Scalable technologies like crossfilter exist. Want to support those in the context of an R environment. Need to support full 2-way communication between computing in the cloud and interaction on the web.
\item Other requirements e.g. Kandel's user study. Not our contribution but things we must address, mentioned in related work section. Heer/Agrawala design considerations for collaborative visual analytics.
\end{itemize}

@@

% limitations
Previous studies have pointed out difficulties in achieving the flexibility,
scalability and maintainability expected of production software with experimental
code created by data analysts.
These properties cannot be ensured by any programming environment,
but suitable tools can help well-motivated analysts and programmers to create
robust applications with much less effort.

>>>>

% \title{RCloud: Integrating Exploratory Visualization, Analysis and Deployment}
\title{RCloud: DevOps for Visual Analysis and Data Science}

@@

%% Abstract section.
\abstract{ Consider the emerging role of a data science team within an
  organization. Individual data scientists and statisticians usually
  work on loosely related problems, and must devise effective ways to 
  share their findings and move results from exploratory data analysis
  into automated diagnostics and reports deployed for wider consumption.
  There are two problems with the current
  practice. First, there are gaps in this workflow: EDA is performed
  with one set of tools, and automated reports and deployments with another.
  Second, EDA environments often assume a single-developer perspective,
  while data scientist teams would get much benefit from easier sharing
  of scripts and data feeds, experiments, annotations, and automated
  recommendations, which are well beyond what traditional version control
  systems provide. \todo{sell idea, not system} In this paper, we propose RCloud, a system that supports
  collaborative data analysis, visualization and web deployment. We will
  discuss the design decisions, tradeoffs and limitations, and compare RCloud
  to other current proposals.
} % end of abstract

@@

% What's the area
Consider the emerging role of a data science team within an
organization today \cite{Keim:2008:VAS}.

% How do people do it today?
Individual data scientists and statisticians usually work on loosely
related problems, and must devise effective ways to share their findings
and move results from exploratory data analysis to automated diagnostics
and reports that are deployed for wider consumption.
They generally rely on a patchwork of resources shared informally,
or at best, through systems that attack individual tasks or functions
in the data analysis workcycle.

% Why does this not work?
There are two problems with the current
practice. First, there are gaps in the workflow:
Exploratory Data Analysis (EDA) is done with one set of tools,
and automated reports and deployments with another. This makes
transferring results from EDA to production slow, expensive, and error-prone.
Second, EDA environments often assume a single-developer perspective,
while data scientist teams could greatly benefit from convenient sharing of
scripts, data sets, data feeds, experiments, annotations, and automated
analyses and recommendations. These features are well beyond what existing
repositories and version control systems provide.

% What are we going to do about this?
\todo{Sell idea, not system} In this paper, we investigate the hypothesis
that the practice of visual analytics can advance by adopting techniques
from information retrieval and collaborative work environments. To investigate
these ideas, we designed and deployed a prototype environment called RCloud,
that supports collaborative data analysis, visualization and web deployment.
We will report on the design decisions, tradeoffs and limitations, and compare
the concepts in RCloud to similar proposals.

\todo{Define the requirements for the system}

We begin by noting the main requirements for a data analysis environment for
collaboration and transfer of results into production.

begin{itemize}
\item Technology transfer. Experiments in notebooks developed by analysts need to become accessible in the production environment, making it easy to push new features, prototypes, etc.
\item Dashboarding. To support the above, there should be a mode in which non-experts
can operate the experiments without coding.
\item Version control as a primary concept in the repository. There needs to be a way to control the visibility of prototypes as they are being developed and tested.
\item Modern interactive technologies. People want to interact with data to explore it. Scalable technologies like crossfilter exist. Want to support those in the context of an R environment. Need to support full 2-way communication between computing in the cloud and interaction on the web.
\item Other requirements e.g. Kandel's user study. Not our contribution but things we must address, mentioned in related work section. Heer/Agrawala design considerations for collaborative visual analytics.
\end{itemize}

@@

% limitations
Previous studies have pointed out difficulties in achieving the flexibility,
scalability and maintainability expected of production software with experimental
code created by data analysts.
Although these properties cannot be enforced by any programming environment,
suitable tools can help well-motivated analysts and programmers to create
robust applications with much less effort.
